\ieeetitle

### 2017-01-12 report/definitions.tex
General comments.

\section{Definitions}
\label{sec:definitions}
A process or peer is defined as an actor in our system running the application that needs total order. Each process communicates with other processes in the distributed system, exchange events, and order them together. \hm{I changed it, but you switched from the present to the future tense in the same sentence. You might as well check the rest of the document for other instances, if any.} \jt{I checked the whole document again I think it is fixed now}
\par
An event $j$ is defined as: $$e_j \triangleq (broadcasterID,timestamp,payload)$$ where the timestamp is the local time at the process when the event is created and is used with the broadcasterID to order events. The payload represents the data sent. %\hm{Not clear. Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
\par
A ball is an abstraction taken from the balls-and-bins problem. In practice, we bundle sets of events together and send them as one packet to reduce the network traffic overhead.

>>>>

\section{Definitions}
\label{sec:definitions}
\mm{merge part about event with background section. The part about the conventions/settings used in the experiments should be moved to the evaluation.}

A process or peer is defined as an actor in our system running the application that needs total order. Each process communicates with other processes in the distributed system, exchange events, and order them together. \hm{I changed it, but you switched from the present to the future tense in the same sentence. You might as well check the rest of the document for other instances, if any.} \jt{I checked the whole document again I think it is fixed now}
\par
An event $j$ is defined as: $$e_j \triangleq (broadcasterID,timestamp,payload)$$ where the timestamp is the local time at the process when the event is created and is used with the broadcasterID to order events. The payload represents the data sent. %\hm{Not clear. Define it mathematically with something like $e_k=(id,time,payload)$ where the time is the (local) time at the process when the event is created.}
\par
A ball is an abstraction taken from the balls-and-bins problem. In practice, we bundle sets of events together and send them as one packet to reduce the network traffic overhead.

### 2017-01-12 report/evaluation.tex
General comments.

\section{Evaluation}
\label{sec:evaluation}
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
%\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
In this section we present our results. In the following benchmarks \epto uses a constant $c$ to set the probability of a hole appearing. We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions. We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
\par 
We used specific settings recommended by \jgroups for a big cluster.
\par
Every \jgroups test run with churn is run once killing the coordinator and once not killing it. We use the following nomenclature to differentiate both tests:
\begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
	\item[\textbf{\jgroups-coord}:] The coordinator is killed once.
	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive throughout the experiment.
\end{description}
\par
\textbf{Testbed.} Our work uses the same cluster first used in \autocite{vaucher2016erasure}. It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has access 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. \eptotester is packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
\par
\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global event throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
\par
We compare \epto and \jgroups based on their bandwidth, the local and global time to deliver every expected events, their local dissemination stretch and finally the number of events sent on average.
\subsection{Bandwidth}
The initial bandwidth peaks observed for \epto are due to the PSS initialization. When a node starts, it receives an initial view and then quickly runs the PSS active thread four times to jump-start itself.
 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/bandwidth-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Bandwidth percentiles of nodes during a stable experiment}
 	\vspace{-2mm} 
 	\label{fig:bandwidth}
 \end{figure}
In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups. It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}. However, in \jgroups most of the work is done solely by the coordinator. We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.6}{\mbps}.

@@

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups. Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}. Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.
 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/bandwidth-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Bandwidth percentiles of nodes during an unstable experiment with synthetic churn}
 	\vspace{-2mm} 
 	\label{fig:bandwidth-churn}
 \end{figure}
\par
In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns. In the first one we kill one node per minute. In the second one, we still kill one node per minute, but we immediately create a new one. For \jgroups we run the benchmarks once without killing the coordinator and once killing it.

@@

\subsection{Total GigaBytes sent/received}
In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.
\par
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received in a stable system}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{llSSS}
		\toprule
		&& \multicolumn{3}{c}{Cluster parameters} \\
		\cmidrule{3-5}
		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
						    &{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
		\midrule
		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
							   &{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth} 
\end{table}
In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups. This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect. Here we are well within this ratio.
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received with a synthetic churn}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{lSSS}
		\toprule
		&& \multicolumn{2}{c}{Churn parameters} \\
		\cmidrule{3-4}
		{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
							&{Sending}& 21.21(025) & 26.57(032)\\
		\midrule
		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
		\midrule
		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth-churn} 
\end{table}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a \jgroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times}
 	\vspace{-2mm}
 	\label{fig:local-times} 
 \end{figure*}

 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times with churn}
 	\vspace{-2mm} 
 	\label{fig:local-times-churn} 
 \end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.
\subsection{Global Times}
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/global-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Global dissemination times}
 	\vspace{-2mm}
 	\label{fig:global-times}  
 \end{figure*}

@@

>>>>

\section{Evaluation}
\label{sec:evaluation}
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth.csv}\tableaverage
%\pgfplotstableread[col sep=comma]{csv-data/average-bandwidth-real-trace.csv}\tableaveragereal
%\pgfplotstableread[col sep=comma]{csv-data/plot-global-time-cdf.csv}\tableglobaltime
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-deltas-cdf.csv}\tablelocaldeltas
%\pgfplotstableread[col sep=comma]{csv-data/plot-local-time-cdf.csv}\tablelocaltime
In this section we present our results. In the following benchmarks \epto uses a constant $c$ to set the probability of a hole appearing. We set $c = 4$ in all our benchmarks so as to not experience any holes as \jgroups does not produce holes under normal conditions. We use a $\delta$ period of \SI{100}{\milli\second} for tests with no churn or synthethic churn and \SI{250}{\milli\second} for tests following a real trace.
We used specific settings recommended by \jgroups for a big cluster.\mm{space allowing, put the configuration used in the paper, otherwise point to the git file.}
Every \jgroups test run with churn is run once killing the coordinator and once not killing it. We use the following nomenclature to differentiate both tests:
\begin{description}[\IEEEsetlabelwidth{\jgroups-nocoord: }]
	\item[\textbf{\jgroups-coord}:] The coordinator is killed once.
	\item[\textbf{\jgroups-nocoord}:] The coordinator is purposely kept alive throughout the experiment.
\end{description}
\textbf{Testbed.} \mm{this sentence is not necessary}Our work uses the same cluster first used in \autocite{vaucher2016erasure}. It consists of 20 machines interconnected by a \SI{1}{\gbps} switched network. Each machine has an 8-core Intel Xeon CPU and \SI{8}{\giga\byte} of RAM. We deploy 12 virtual machines on these hosts. Each virtual machine has access 4 VCPUs as well as \SI{7}{\giga\byte} of RAM. We use KVM as our hypervisor. Each VM uses Debian as its O/S. \eptotester is packaged as docker images. We use Docker 1.12 and its new functionality Docker Swarm to orchestrate our services. Each docker container has a memory restriction of \SI{300}{\mega\byte} of RAM to be able to run them all together.
\textbf{Experiment parameters.} Every benchmark except the one following a real trace are run 10 times, each during \SI{20}{\minute}. When there is synthetic churn, the churn starts \SI{30}{\second} after the benchmark and run for \SI{17}{\minute}. The benchmarks following a real trace are run 5 times. The trace is speed-up 2x, which means we follow \SI{2}{\hour} worth of trace in \SI{1}{\hour}. Every benchmark run with churn is run with the $(100,50)$ parameters. When we have a global event throughput lower than the number of actual peers, we use a uniform random number generator to artificially reduce the global event throughput. For example, using $(100,50)$ parameters, each peer has a 50\% probability of sending an event at each second.
We compare \epto and \jgroups based on their bandwidth, the local and global time to deliver every expected events, their local dissemination stretch and finally the number of events sent on average.
\subsection{Bandwidth}
The initial bandwidth peaks observed for \epto are due to the PSS initialization. When a node starts, it receives an initial view and then quickly runs the PSS active thread four times to jump-start itself.
 	\includegraphics[width=\linewidth]{figures/bandwidth-nochurn.pdf}
 	\caption{Bandwidth percentiles of nodes during a stable experiment}
 	\label{fig:bandwidth}
In \autoref{fig:bandwidth} we observe that \epto has a worse baseline compared to \jgroups. It uses a median bandwidth of approximately \SI{1}{\mbps} for $(50,50)$ whereas \jgroups uses a median bandwidth of less than \SI{0.2}{\mbps}. However, in \jgroups most of the work is done solely by the coordinator. We can clearly see this as the 100th percentile is much higher than the rest and uses approximately \SI{.6}{\mbps}.

@@

Comparing \epto and \jgroups in terms of bandwidth when we increase the number of peers, \epto scales better than \jgroups. Where \jgroups basically has to double the bandwidth usage of the coordinator, \epto only increases it marginally \jt{logarithmic I think as was expected}. Thus in a scenario where we have many peers \epto will be more efficient than \jgroups at not reaching the maximum available bandwidth on a given node.\mm{it would be perfect if we could experimentally show this.}
 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/bandwidth-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Bandwidth percentiles of nodes during an unstable experiment with synthetic churn}
 	\vspace{-2mm} 
 	\label{fig:bandwidth-churn}
 \end{figure}
\par
In \autoref{fig:bandwidth-churn} We analyze two different synthetic churns. In the first one we kill one node per minute. In the second one, we still kill one node per minute, but we immediately create a new one. For \jgroups we run the benchmarks once without killing the coordinator and once killing it.

@@

\subsection{Total GigaBytes sent/received}
In all cases, \jgroups receives more bytes than it sends. This is due to the fact that we do not measure the bandwidth on the TCPGOSSIP, which is in charge of view changes.\mm{we need to explain why this is the case}
\par
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received in a stable system}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{llSSS}
		\toprule
		&& \multicolumn{3}{c}{Cluster parameters} \\
		\cmidrule{3-5}
		{Protocol}&& {$(50,50)$} & {$(50,100)$} & {$(100,50)$} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 10.84(016) & 22.31(039) & 26.01(027) \\
						    &{Sending}& 10.84(016) & 22.31(039) & 26.01(027)\\
		\midrule
		\multirow{2}{*}{\jgroups}&{Receive}& 0.78(003) & 1.45(001) & 1.88(001)\\
							   &{Sending}& 0.77(003) & 1.44(001) & 1.84(001)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth} 
\end{table}
In \autoref{table:total-bandwidth}, \epto has a worse baseline than \jgroups. This is expected as \epto sends $c*n*\log_2 n$ messages per events and \jgroups sends at least $n$ messages per event so we should have at least $c*\log_2 n$ more messages sent in \epto if \jgroups is perfect. Here we are well within this ratio.
\begin{table}[hpt]
	\centering
	\caption{Total \si{\giga\byte} sent/received with a synthetic churn}
	\sisetup{table-format=2.2, separate-uncertainty, table-figures-uncertainty = 2, table-align-uncertainty}
	\resizebox{\linewidth}{!}{
	\begin{tabular}{lSSS}
		\toprule
		&& \multicolumn{2}{c}{Churn parameters} \\
		\cmidrule{3-4}
		{Protocol}&& {1 kill/minute} & {1\{kill,add\}/minute} \\
		\midrule
		\multirow{2}{*}{\epto}&{Receive}& 21.00(024) & 26.32(032)\\
							&{Sending}& 21.21(025) & 26.57(032)\\
		\midrule
		\multirow{2}{*}{\jgroups-coord}&{Receive}& 1.47(002) & 1.75(002)\\&{Sending}& 1.43(002) & 1.70(002)\\
		\midrule
		\multirow{2}{*}{\jgroups-nocoord}&{Receive}& 1.45(001) & 1.73(002)\\&{Sending}& 1.41(001) & 1.68(002)\\
		\bottomrule
	\end{tabular}
	}
	\label{table:total-bandwidth-churn} 
\end{table}
In \autoref{table:total-bandwidth-churn} we see that \jgroups total bandwidth usage is smaller when there is churn. One hypothesis for this is that a \jgroups replica takes a longer time to start up compared to stopping a replica. Therefore the overall benchmark has a longer time with less than 100 replicas. We also do not see a difference whether we kill the coordinator or not. This can be explained by the fact that before the detection of a faulty coordinator \jgroups is forced to a halt for period of up to \SI{20}{\second}. The big spike afterwards compensates for this hole.
\subsection{Local Times}
\label{sub:local-times}
\mm{explain what this is measuring, same for the other metrics}
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times}
 	\vspace{-2mm}
 	\label{fig:local-times} 
 \end{figure*}

 \begin{figure}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/local-times-synth-churn.pdf}
 	\vspace{-2mm} 
 	\caption{Local dissemination times with churn}
 	\vspace{-2mm} 
 	\label{fig:local-times-churn} 
 \end{figure}
In \autoref{fig:local-times}, \jgroups delivers all events quicker than \epto in all scenarios, even when churn is involved as is shown in \autoref{fig:local-times-churn}. However, \epto is not too far behind. The difference between \epto and \jgroups is likely to be even smaller when running them in a real WAN network due to the latency. \epto in our configuration has a $\delta$ period of \SI{100}{\milli\second} and is thus handicapped against \jgroups in a LAN environment, because it only increments the TTL of an event every \SI{100}{\milli\second}. We expect \epto to outperform \jgroups in this situation when we have a high number of peers in the system or when the network has a higher latency.\mm{we could try to setup an AWS test in mutliple availability zones to see what happens.}
\subsection{Global Times}
 \begin{figure*}[hpt]
 	\centering
 	\includegraphics[width=\linewidth]{figures/global-times-nochurn.pdf}
 	\vspace{-2mm} 
 	\caption{Global dissemination times}
 	\vspace{-2mm}
 	\label{fig:global-times}  
 \end{figure*}

@@

\mm{the part regarding  could probably be expanded and moved to the previous section, where we discuss the implementation. the part regarding Jgroups could also be expanded. do we have a rough idea of the frequency of this ? it's also important do discuss how we measure holes so it's clear that this is not a problem on our side }

### 2017-01-12 report/framework.tex
General comments.

@@

We implement the payload as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port.  We compare \epto to the deterministic total order algorithm SEQUENCER provided by \jgroups. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker does not support MULTICAST. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.

>>>>

\mm{along the lines of presenting the tool as a contribution, could this be more elaborated ? The part that briefly describes EpTO could be moved to background. The part about the use of UDP, tracker et all belongs to the implementation. }

@@

We implement the payload as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port.  We compare \epto to the deterministic total order algorithm SEQUENCER provided by \jgroups. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers as Docker does not support MULTICAST. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.\mm{cite to the jgroups documentation}

### 2017-01-12 report/introduction.tex
General comments.

@@

Until now, the creators of \epto have only tested this algorithm in a simulated environment. In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems. We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments. \epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view. These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm. Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases} in the system.

>>>>

\mm{Reorder a little bit such that the motivation and existing approaches to total order are introduced first and then epto is presented as an alternative approach. Also stress the motivation is to run a TO algorithm in a very large scale and draw insights from it, similarly to what was done in the papers "paxos made live" and "worldwide consensus.". Point out that there are few works evaluating such protocols in a really large scale.}\mm{We also need to make the contribution clear: In this case it is an implementation of a total order algorithm and making it available as open source. It would be good if we could put this on a more generic light, along the lines of "a set of tools to evaluate large scale algorithms" or at least present the tool itself, eptotester as a contribution, wdyt ? this would make the paper stronger }

@@

Until now, the creators of \epto have only tested this algorithm in a simulated environment. In this work, we implement \epto in pure Kotlin\footnote{\href{https://kotlinlang.org/}{https://kotlinlang.org/}} and introduce a benchmarking solution to show that \epto is suitable for real-world large-scale distributed systems. We compare \epto to the deterministic total order algorithm provided by \jgroups \autocite{jgroups} in both stable and unstable environments. \epto uses a PSS CYCLON \autocite{Voulgaris2005} to obtain a random view, coupled with a tracker that keeps track of dead and alive peers to obtain a first initial view.\mm{this is imprecise and can lead to confusion, rephrase} These benchmarks can easily be launched and scaled on a cluster using Docker and Docker Swarm. Furthermore, we can inject synthetic or real churn\footnote{Using the Failure Trace Archive databases}\mm{link} in the system.

### 2017-01-12 report/ordering.tex
General comments.

\section{Ordering Algorithms}
\label{sec:ordering}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\par
In other words, total order is an ordering that defines the exact order of every event in the system. On the other hand, partial ordering only defines the order between certain key events that depend on each other. Partial order can be useful since it is less costly to implement. However, in some cases the order of all events is important. For example, imagine we have multiple databases around the world. We want them to appear as if it is only one database. To achieve this, every operation done a particular database would have to be replicated on all the others, we would then have to make sure that every operation is executed in the same order on every database to end up in the same state. \hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}\jt{I came up with a possible total order application. However I am not sure if it is the best use case. Maybe Miguel has a better idea.}

>>>>

\section{Ordering Algorithms}
\label{sec:ordering}
\mm{rename to background. we also need to include some related work on TO (see epto paper) and on practical evaluation (e.g.: the papers pointed above plus splay}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

\subsection{Total Order Algorithms}
A totally ordered set of events is a partially ordered set which satisfies one additional property:
\begin{itemize}
	\item Totality (trichotomy law): For any $a, b \in S$, either $a \leq b$  or $b \leq a$.
\end{itemize}
\par
In other words, total order is an ordering that defines the exact order of every event in the system. On the other hand, partial ordering only defines the order between certain key events that depend on each other. Partial order can be useful since it is less costly to implement. However, in some cases the order of all events is important. For example, imagine we have multiple databases around the world. We want them to appear as if it is only one database. To achieve this, every operation done a particular database would have to be replicated on all the others, we would then have to make sure that every operation is executed in the same order on every database to end up in the same state. \hm{It would be enlightening to be more precise here and add a large-scale distributed application for which total order is important.}\jt{I came up with a possible total order application. However I am not sure if it is the best use case. Maybe Miguel has a better idea.}\mm{we can point that that TO is equivalent to consensus and used in approaches such as distirbuted state machines. Actually it is the fact that TO is not scalable that lead to the emergence of all the work on eventual consistency.}
